% \chapter{Tài liệu tham khảo}
% \addcontentsline{toc}{chapter}{Tài liệu tham khảo}
% \label{ch:references}
\newpage
\renewcommand{\bibname}{Tài liệu tham khảo}
\begin{thebibliography}{99}

\bibitem{chen2017eyeriss} Chen, Y.-H., Krishna, T., Emer, J. S., \& Sze, V. (2017). Eyeriss: An Energy-Efficient Reconfigurable Accelerator for Deep Convolutional Neural Networks. \textit{IEEE Journal of Solid-State Circuits}, 52(1), 127–138. 

\bibitem{li2025pixel} Li, Z., Zhang, Z.s, Hu, J., Meng, Q., Shi, X., Luo, J., Wang, H., Huang, Q., \& Chang, S. (2025). A High-Performance Pixel-Level Fully Pipelined Hardware Accelerator for Neural Networks. \textit{IEEE Transactions on Neural Networks and Learning Systems}, 36(5), 7970–7981. 

\bibitem{zhang2017opencl} Zhang, J., \& Li, J. (2017). Improving the Performance of OpenCL-based FPGA Accelerator for Convolutional Neural Network. In \textit{Proceedings of the 2017 ACM/SIGDA International Symposium on Field-Programmable Gate Arrays (FPGA '17)}, 25–34. 

\bibitem{liu2020riscv} Liu, Z., Jiang, J., Lei, G., Chen, K., Qin, B., \& Zhao, X. (2020). A Heterogeneous Processor Design for CNN-Based AI Applications on IoT Devices. \textit{Procedia Computer Science}, 174, 2–8. 
\bibitem{du2017rsa} Du, L., Du, Y., Li, Y., \& Chang, M.-C. F. (2017). A Reconfigurable Streaming Deep Convolutional Neural Network Accelerator for Internet of Things. \textit{IEEE Transactions on Circuits and Systems I: Regular Papers}, 65(1), 198–208.


\bibitem{chu2017fpga} Chu, P. P. (2017). \textit{FPGA Prototyping by SystemVerilog Examples: Xilinx MicroBlaze MCS SoC Edition}. Wiley.

\bibitem{brown2013fundamentals} Brown, S., \& Vranesic, Z. (2013). \textit{Fundamentals of Digital Logic with Verilog Design} (3rd ed.). McGraw-Hill Education.

\bibitem{arm2022axi} Arm Limited. (2022). \textit{Introduction to AMBA AXI}. [Online]. Available: https://developer.arm.com

\bibitem{arm2010axistream} Arm Limited. (2010). \textit{AMBA AXI-Stream Protocol Specification} (IHI0051B).

\bibitem{xilinx2012axi} Xilinx, Inc. (2012). \textit{LogiCORE IP AXI Interconnect (v1.03.a) Data Sheet} (DS768).

\bibitem{pena2020uart} Peňa, E., \& Legaspi, M. G. (2020). \textit{UART: A Hardware Communication Protocol Understanding Universal Asynchronous Receiver/Transmitter}. Analog Dialogue, Vol. 54. [Online]. Available: https://www.analog.com

\bibitem{dhaker2018spi} Dhaker, P. (2018). \textit{Introduction to SPI Interface}. Analog Dialogue, Vol. 52. [Online]. Available: https://www.analog.com

\bibitem{legaspi2021i2c} Legaspi, M. G., \& Peňa, E. (2021). \textit{I2C Communication Protocol: Understanding I2C Primer, PMBus, and SMBus}. Analog Dialogue, Vol. 55. [Online]. Available: https://www.analog.com

\end{thebibliography}